---
title: "tidy_thursday_8_5_2021"
output: html_document
---

# August meeting of the South Coast MA UseR Group: `DataExplorer` Package
This document tests the features of the `DataExplorer` R package by 1) working through examples in the package documentation, and 2) using these features to explore a Tidy-Tuesday dataset. 

Start by installing and loading the package:
```{r}
# install.packages("DataExplorer")
library(DataExplorer)
```

Load other useful packages
```{r}
library(nycflights13) # Data
library(tidyverse) 
```

## Work through the package documentation
This document walks through `DataExplorer` package features, borrowing from the following:

- The package introduction: http://boxuancui.github.io/DataExplorer/
- The package vignettes: https://boxuancui.github.io/DataExplorer/articles/dataexplorer-intro.html
- Whenever possible the following two datasets were used in the examples (rather than switching between many datasets as in the documentation):
 - `airquality` from the `DataExplorer` package
 - `flights` from the `nycflights13` package


Create a report that applies all the major features of this package & returns an html report
```{r, include=FALSE}
create_report(airquality)
```

### The report breakdown
#### Describe data
View a description of the data, including structure, size, missing information:
```{r}
introduce(airquality)
```

Plot this introductory data:
```{r}
plot_intro(airquality)
```

Plot the distribution of missing values by variable
```{r}
plot_missing(airquality)
```

Examine the data structure
```{r}
plot_str(airquality)
```

This will also break down lists of data nicely
```{r}
listdata <- list(airquality, flights)
plot_str(test_data)
```

Optional: the orientation of this structure can be switched to a circle
```{r}
plot_str(test_data, type="r")
```

#### Continuous variables
Plot histogram of continuous variables 
```{r}
plot_histogram(airquality)
```

Optional: add plot of density distribution for continous variables
```{r}
plot_density(airquality)
```

QQ plots are automatically considered for continuous variables:
```{r}
plot_qq(airquality)
```

Optional: also break down qqplot by variable
```{r}
plot_qq(airquality, by = "Month")
```

Optional: Plot boxplot of continuous variables relative to another continuous or discrete variable
```{r}
# Continuous variable
plot_boxplot(airquality, by = "Ozone")
# Discrete variable
plot_boxplot(flights, by = "origin")
```

Optional: Plot scatterplot of continuous variables relative to another continuous variable
```{r}
plot_scatterplot(airquality, by = "Ozone")
```


#### Discrete variables
Plot frequencies for discrete variables in the dataset (none in `airquality` so use `flights` dataset instead)
```{r}
plot_bar(flights)
```

Optional: plot distribution of discrete variables relative to a continuous variable (e.g. departure delay)
```{r}
plot_bar(flights, with = "dep_delay")
```

Optional: plot frequency distribution of discrete variables colored relative to a discrete variable (e.g. departure delay)
```{r}
plot_bar(flights, by = "origin")
```


#### All variables
Plot correlation heatmap
```{r}
plot_correlation(airquality)
```

Missing values will result in no correlation calculation, to resolve remove rows with NAs:
```{r}
airquality[complete.cases(airquality),] %>% plot_correlation()
```

The `type` argument can be used to show only continuous variables (type = "c") or discrete variables (type = "d")
```{r}
flights[complete.cases(flights),] %>% plot_correlation(type = "c") # Only 1 year in data
flights[complete.cases(flights),] %>% plot_correlation(type = "d")
```


Run and visulaize principal component analysis (must remove rows containing NAs in order to run)
```{r}
plot_prcomp(na.omit(airquality))
```


### Feature engineering
Create new features using existing data!

Replace missing values in flights dataset, 0s for continuous variables, "unknown" for discrete variables
```{r}
noMissing <- set_missing(flights, list(0, "unknown")) 
noMissing
plot_missing(noMissing) # No longer has missing values
```

Group sparse data into a single category based on threshold (bottom 20% of counts), update = TRUE updates original data
```{r}
temp <- group_category(flights, feature = "carrier", threshold = 0.2, update = TRUE)
plot_bar(temp$carrier)
```
May also group by another variable (e.g. carrier grouped by lowest 20% distance traveled)
```{r}
temp <- group_category(flights, feature = "carrier", threshold = 0.2, update = TRUE, measure = "distance")
plot_bar(temp$carrier)
```

Transform discrete variables into a binary format with `dummify`

Remove data columns (e.g. one with many missing values or not of interest) using `drop_columns`, performs the opposite operation of `dplyr::select` 
E.g. to remove the Ozone column run:
```{r}
drop_columns(airquality, "Ozone")
```

`update_columns` can be used to perform a calculation or transformation on a single variable, similar to `dplyr::mutate` but automatically keeps the same column name for the manipulated variable
E.g. turn Month into a categorical variable:
```{r}
update_columns(airquality, "Month", as.factor)
```

### Data reporting
`create_report()` automatically structures report with many of the plots discussed above based on the data structure (i.e. will not include plots of discrete variables if none present in dataset)

Optional: adding a `y` argument will result in bivariate plots to be added to report (e.g. boxplots and scatterplots)
```{r, include=FALSE}
create_report(airquality, y = "Ozone")
```

Optional: `configure_report()` lets you add/remove plots to customize the report (e.g. scatterplots)
```{r, include=FALSE}
config <- configure_report(add_plot_density = TRUE,
                           global_ggtheme = quote(theme_dark()))

create_report(airquality, config = config)
```


## Will it work for your data? A Tidy-Tuesday test
Try generating a DataExplorer report for a messier Tidy-Tuesday dataset (find the datset here: https://github.com/rfordatascience/tidytuesday)

Load the dataset
```{r, include=FALSE}
tuesdata <- tidytuesdayR::tt_load('2021-08-03')
```

Must have rows & columns
```{r, include = FALSE}
config <- configure_report(add_plot_bar = TRUE, 
                           plot_bar_args = list(by = "gender"))

create_report(tuesdata$athletes, config = config, y = "year")
# Look at country, gender, type over years in scatterplots
```

Plot default barplot
```{r}
plot_bar(tuesdata$athletes)
```