---
title: "MarchUseRMeeting_gganimate"
output: html_document
---

This document contains code to use gganimate, as part of March South Coast MA UseR Group meeting, 2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse) # general use ----
library(gganimate) # to create animated gif ----
library(extrafont) # access and enable fonts ----
library(ggtext)    # style text in ggplot ----
library(glue)      # paste strings ----  
```

## Including Plots

You can also embed plots, for example:

```{r}
us_states <- ggplot2::map_data("state")
residence <- tribble(
  ~city,           ~state,  ~lat,   ~long, ~years, ~description,
  "Raleigh",         "NC", 35.82,  -78.66,  17,    "Childhood",
  "Greenville",      "NC", 35.60,  -77.37,   4,    "Undergrad at ECU",
  "Atlanta",         "GA", 33.76,  -84.42,  10,    "Grad school at Emory<br>Statistician at CDC<br>Lecturer at Emory",
  "San Luis Obispo", "CA", 35.28, -120.66,   3,    "Asst. Professor at Cal Poly SLO",
  "Williamsburg",    "VA", 37.27,  -76.71, 0.5,    "Time with family",
  "Doylestown",      "PA", 40.31,  -75.13,   2,    "Statistician at Adelphi Research"
) 

residence_connections_prelim <- residence %>% 
  mutate(
    # need this to create transition state ----
    city_order = row_number() + 1,
    # where I moved to next, for curved arrows ----
    lat_next = lead(lat),
    long_next = lead(long),
    # label to show in plot, styled using ggtext ---
    label = glue::glue("**{city}, {state}** ({years} yrs)<br>*{description}*"),
    # label of next location ----
    label_next = lead(label)
  ) 

residence_connections <- residence_connections_prelim %>%
  # get first row of residence ----
  slice(1) %>% 
  # manually modify for plotting ----
  mutate(
    city_order = 1,
    label_next = label,
    lat_next = lat,
    long_next = long,
    ) %>% 
  # combine with all other residences ----
  bind_rows(residence_connections_prelim) %>% 
  # last (7th) row irrelevant ----
  slice(1:6) %>% 
  # keep what we neeed ----
  dplyr::select(city_order, lat, long, lat_next, long_next, label_next)


base_map <- ggplot() +
  # plot states ----
  geom_polygon(
    data = us_states,
    aes(
      x     = long, 
      y     = lat, 
      group = group
      ),
    fill  = "#F2F2F2",
    color = "white"
  ) +
  # lines for pins ----
  geom_segment(
    data = residence,
    aes(
      x    = long,
      xend = long,
      y    = lat,
      yend = lat + 0.5
      ),
    color = "#181818",
    size = 0.3
    ) +
    # pin heads, a bit above actual location, color with R ladies lighter purple ----
  geom_point(
    data = residence,
    aes(
      x = long, 
      y = lat + 0.5
      ),
    size = 0.5,
    color = "#88398A"
  ) +
  theme_void()
base_map
```

Animate
```{r}
anim <- base_map +
  # show arrows connecting residences ----
  geom_curve(
    # do not include 1st residence in arrows as no arrow is intended ----
    # and inclusion messes up transition ---
    data = residence_connections %>% slice(-1),
    # add slight adjustment to arrow positioning ----
    aes(
      y     = lat - 0.1,
      x     = long,
      yend  = lat_next - 0.2,
      xend  = long_next,
      # group is used to create the transition ----
      group = seq_along(city_order)
    ),
    color = "#181818",
    curvature = -0.5,
    arrow = arrow(length = unit(0.02, "npc")),
    size  = 0.2
  ) +
  # add in labels for pins, with inward positioning ----
  # show labels either top left or top right of pin ----
  geom_richtext(
    data = residence_connections,
    aes(
      x     = ifelse(long_next < -100, long_next + 1, long_next - 1),
      y     = lat_next + 5,
      label = label_next,
      vjust = "top",
      hjust = ifelse(long_next < -100, 0, 1),
      # group is used to create the transition ----
      group = seq_along(city_order)
    ),
    size = 2,
    label.colour = "white",
    # R ladies purple ----
    color = "#562457",
    # R ladies font used in xaringan theme ----
    family = "Lato"
  ) +
  # title determined by group value in transition ----
  ggtitle("Home {closest_state} of 6") +
  # create animation ----
  transition_states(
    city_order,
    transition_length = 2,
    state_length = 5
    ) +
  # style title ----
  theme(
    plot.title = element_text(
      color = "#562457",
      family = "Permanent Marker",
      size = 12
      )
    )
# render and save transition ----
# the default nframes 100 frames, 150 makes the gif a bit longer for readability ----
# changing dimensions for output w/ height & width ----
# increasing resolution with res ----
animate(anim, nframes = 150, height = 2, width = 3, units = "in", res = 150)
```


Shape files: For coastlines/country/state borders I use rnaturalearth
Keeps me from needing to remember file names for different .shp files

Here are the two lines to get global country polygons:
Adam Kemberling to Everyone (4:42 PM)
library(rnaturalearth)
world  <- ne_countries() %>% st_as_sf(crs = 4326)


For saving/displaying annimation: gifski
https://ropensci.org/blog/2018/07/23/gifski-release/

```{r}
library(marmap)
Bathy <- getNOAA.bathy(lon1 = -75, lon2 = -62,
                       lat1 = 39, lat2 = 48, resolution = 1)
# convert data to matrix:
Bathy_Final <- as.matrix(Bathy)
class(Bathy_Final) <- "matrix"

# now reshape it to long format and plot
BathyData <- Bathy_Final %>%
  as.data.frame() %>%
  rownames_to_column(var = "lon") %>%
  gather(lat, value, -1) %>%
  mutate_all(funs(as.numeric)) 

```

```{r}
library(ggimage)
# Example fish locations
LocDat <- data.frame(
  long = c(-69.5 , -69, -67),
  lat = c(42, 40.7, 41.3),
  imagecol = rep("/Users/ahart2/Library/R/3.6/library/gganimate/help/figures/logo.png",3),
  stringsAsFactors = FALSE,
  position = c(1,2,3)
) 

World <- map_data("world")

AnimateBathy <- ggplot(BathyData) +
  geom_contour(aes(x = lon, y = lat, z = value), 
               bins = 50, colour = "black") + # bins determine # of contours (try 50 vs. 100)
  geom_polygon(data = World, 
               aes(x=long, y = lat, group = group), 
               fill = "grey", 
               color="black") + 
  coord_fixed(xlim = c(-74, -64),  ylim = c(40, 47.3), ratio = 1.2)  +
  theme(panel.background = element_rect(fill = "dodgerblue4"), line = element_blank()) +
  geom_image(data=LocDat, mapping = aes(x = long, y = lat, image = imagecol), size = 0.085) + 
  transition_time(time = position) +   # slides point from first to second position, other transitions available
  labs(title = paste('Position: {frame_time}'))

AnimateBathy
```


Use transition_reveal () to show new info (e.g. display previously revealed data & move to a new point) https://jamesepaterson.github.io/jamespatersonblog/2020-01-02_animatingtrackingdata.html


Transitions in cheat sheet: https://ugoproto.github.io/ugo_r_doc/pdf/gganimate.pdf
