---
title: "R Notebook"
output: html_notebook
---

# February 2022 UseR meeting ggvis package

Load packages
```{r}
library(dplyr)
library(ggvis)
```
## Tutorial walk-through
```{r}
mtcars %>%
  ggvis(x=~mpg, y=~disp) %>%
  mutate(disp=disp/61.0237) %>% # convert to litres
  layer_points()
```

Add interactivity
```{r}
mtcars %>% 
  ggvis(~wt, ~mpg, 
    size := input_slider(10, 100),
    opacity := input_slider(0, 1)
  ) %>% 
  layer_points()
```

Really cool feature = can include interactive plots in Rmd documents knit to htmls

Add tooltips to highlight values (here weight)
```{r}
mtcars %>% ggvis(~wt, ~mpg) %>% 
  layer_points() %>% 
  add_tooltip(function(df) df$wt)
```


## My explorations
Can you add interactive scale bars that switch between categorical variables? - maybe, still exploring
https://ggvis.rstudio.com/interactivity.html
```{r}
library("palmerpenguins")

penguins %>%
  ggvis(x = ~bill_length_mm, y =~bill_depth_mm) %>%
  layer_points()
```

Useful example for using prop: https://stackoverflow.com/questions/39666748/ggvis-using-a-variable-to-assign-x-property-using-props

Set up shape shift based on selected choice, doesn't correspond to data in any way
```{r}
# p1 <- prop("x", input_select(c("bill_length_mm", "flipper_length_mm", "body_mass_g")))
# p2 <- prop("shape", input_select(choices = c("cross", "diamond", "cross"), map = c("sex", "species", "island")))

p3 <- prop("shape", input_select(choices = c("sex" = "cross", "species" = "diamond"))) # read help document for input_select

#p4 <- prop("shape", input_select(choices = c("sex" , "species"))) # read help document for input_select

head(penguins, n=20) %>%
  ggvis(x=~bill_length_mm, y =~bill_depth_mm, shape := p3) %>%
  layer_points() 
  
```

Plot with categorical factor for shape
```{r}
head(penguins, n=20) %>%
  ggvis(x=~bill_length_mm, y =~bill_depth_mm, shape =~factor(sex)) %>%
  layer_points() 
  
```

Tried switching so plots shape based on categorical factor selected
```{r}
# p1 <- prop("x", input_select(c("bill_length_mm", "flipper_length_mm", "body_mass_g")))
# p2 <- prop("shape", input_select(choices = c("cross", "diamond", "cross"), map = c("sex", "species", "island")))

# p3 <- prop("shape", input_select(choices = c("sex" = "cross", "species" = "diamond"))) # read help document for input_select

p4 <- prop("shape", input_select(choices = c("sex" = ~factor(sex) , "species" = ~factor(species)), map = identity)) # read help document for input_select

# head(penguins, n=20) %>%
#  ggvis(x=~bill_length_mm, y =~bill_depth_mm, shape = ~factor(sex)) %>%
#  layer_points() 

head(penguins, n=20) %>%
  ggvis(x=~bill_length_mm, y =~bill_depth_mm, shape := p4) %>%
  layer_points() 
  
```




