---
title: "tidytuesday_12-3_2020"
output: html_document
---

# Tidy Tuesday South Coast MA UseR Group December Meeting
```{r setup, include=FALSE}
library(tidyverse)
```


```{r}
library("tidytuesdayR")
tuesdata <- tidytuesdayR::tt_load('2020-12-01')
data <- tuesdata$shelters
```
# Explore data
```{r}
summarize(data)
summary(data)
glimpse(data)
colnames(data)
```

# Switch gears entirely to look at colors
Where to find color gradients/scales:

- Colorbrewer 
- viridis package & related color schemes

The scales package lets you print colors to screen using names:
```{r}
library(scales)
show_col(c("#9C89B8", "#F0A6CA", "blue"))
```

# Try making a map of shelters with a google map background
```{r}
library(lubridate)
library(data.table)
library(ggrepel)
library(tidyverse)
```

First tried to set up map with goolge maps background using ggmap package. I found this article (https://www.r-graph-gallery.com/324-map-background-with-the-ggmap-library.html) and the default lat/long dimensions a good place to start (adding google map background requires link to API that I didn't have time to do).

The maptype gives a couple of themes to work with if you don't use a google map background.
```{r}
test <- get_stamenmap(maptype = "watercolor")
ggmap(test)

```

Another map type!
```{r}
test <- get_stamenmap(maptype = "toner")
ggmap(test)

```

Finally figured out setting the lat/long dimensions correctly to focus on Toronto area.
```{r}
test <- get_stamenmap(bbox = c(bottom = 41, top = 45, right = -77, left = -81), maptype = "watercolor")
ggmap(test)

```

