---
title: "Tidy Tuesday Contribution 4/31/20"
author: "Amanda Hart"
date: "3/31/2020"
output: html_document
---

First read in the data and load the `tidyverse` package:
```{r}
library(tidyverse)

brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv')
beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv')
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv')
beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')

```

# brewing_materials data
Initial data exploration: 
```{r}
# Brewing_materials
colnames(brewing_materials)
head(brewing_materials)
dim(brewing_materials)
```

# beer_taxed data
Initial data exploration: 
```{r}
colnames(beer_taxed)
head(beer_taxed)
dim(beer_taxed)
```

# beer_states data
Initial data exploration: 
```{r}
colnames(beer_states)
head(beer_states)
dim(beer_states)
unique(beer_states$state)
```

Create a map of current production by year:
```{r}
library(ggplot2)
# The following doesn't contain the beer_states data yet
states1 <- map_data("state")

ggplot() +
  geom_polygon(data = states1, 
               aes(x=long, y = lat, group = group), 
               fill = "white", 
               color="black") +
  coord_map()

# # The following doesn't work:
# library(fiftystater)
# states <- sf::st_as_sf(fifty_states, coords = c("long", "lat"))
# #fiftystates <- st_transform(shellfish, "+proj=longlat +ellps=WGS84 +datum=WGS84")
# fiftystates <- states %>% group_by(id, piece) %>% 
#   summarize(do_union = FALSE) %>%
#   st_cast("POLYGON") %>% 
#   ungroup()
# 
# beer_states %>% 
#   merge_element()
#   group_by(year) 
# 
#   ggplot() + 
#   geom_sf(data = fiftystates, fill = barrels, color = "black")
#   geom_polygon(data = states, 
#                aes(x=long, y = lat, group = group), 
#                fill = "white", 
#                color="black") +
#   coord_map()
```

## brewer_size data
Initial data exploration: 
```{r}
colnames(brewer_size)
head(brewer_size)
dim(brewer_size)
```

Plot the number of brewers in each size category over time:
```{r}
# rank brewer size by number
brewer_size <- brewer_size %>% mutate(brewer_size_ordered = case_when(brewer_size == "Zero Barrels" ~  "A", 
                                                       brewer_size == "Under 1 Barrel" ~ "B", 
                                                       brewer_size == "1 to 1,000 Barrels" ~ "C", 
                                                       brewer_size == "1,001 to 7,500 Barrels" ~ "D", 
                                                       brewer_size == "7,501 to 15,000 Barrels" ~ "E", 
                                                       brewer_size == "15,001 to 30,000 Barrels" ~ "F", 
                                                       brewer_size == "30,001 to 60,000 Barrels" ~ "G", 
                                                       brewer_size == "60,001 to 100,000 Barrels" ~ "H", 
                                                       brewer_size == "100,001 to 500,000 Barrels" ~ "I", 
                                                       brewer_size == "500,001 to 1,000,000 Barrels" ~ "J", 
                                                       brewer_size == "1,000,001 to 6,000,000 Barrels" ~ "K",
                                                       brewer_size == "6,000,001 Barrels and Over" ~ "L",
                                                       brewer_size == "1,000,000 to 6,000,000 Barrels" ~ "M",
                                                       brewer_size == "2,000,000 to 6,000,000 Barrels" ~ "N",
                                                       brewer_size == "1,00,001 to 1,999,999 Barrels" ~ "O",
                                                       brewer_size == "Total" ~ "P"))
brewer_size
```
Create gghighlight plot of # brewers by size over time

```{r}
# Plot number of brewers by size category over time
brewer_size %>% 
  filter(brewer_size_ordered != "M",
         brewer_size_ordered != "N",
         brewer_size_ordered != "O",
         brewer_size_ordered != "NA") %>%
  group_by(year) %>%
  ggplot() +
  geom_line(aes(y = n_of_brewers, x = year, col=brewer_size_ordered)) +
  scale_colour_discrete(name = "Brewer size", labels = c("Zero Barrels", "Under 1 Barrel", "1 to 1,000 Barrels", "1,001 to 7,500 Barrels", "7,501 to 15,000 Barrels", "15,001 to 30,000 Barrels", "30,001 to 60,000 Barrels", "60,001 to 100,000 Barrels", "100,001 to 500,000 Barrels", "500,001 to 1,000,000 Barrels", "1,000,001 to 6,000,000 Barrels", "6,000,001 Barrels and Over", "Total"))

# Plot production by size category over time
brewer_size %>% 
  group_by(year) %>%
  ggplot() +
  geom_line(aes(y = total_barrels, x = year, col=brewer_size_ordered)) +
  scale_colour_discrete(name = "Brewer size", labels = c("Zero Barrels", "Under 1 Barrel", "1 to 1,000 Barrels", "1,001 to 7,500 Barrels", "7,501 to 15,000 Barrels", "15,001 to 30,000 Barrels", "30,001 to 60,000 Barrels", "60,001 to 100,000 Barrels", "100,001 to 500,000 Barrels", "500,001 to 1,000,000 Barrels", "1,000,001 to 6,000,000 Barrels", "6,000,001 Barrels and Over", "Total"))

# 
library(gghighlight)
brewer_size %>%
  group_by(year) %>% filter(total_barrels < 7000000) %>%
    ggplot() +
  geom_line(aes(y = total_barrels, x = year, col=brewer_size_ordered)) +
  scale_colour_discrete(name = "Brewer size", labels = c("Zero Barrels", "Under 1 Barrel", "1 to 1,000 Barrels", "1,001 to 7,500 Barrels", "7,501 to 15,000 Barrels", "15,001 to 30,000 Barrels", "30,001 to 60,000 Barrels", "60,001 to 100,000 Barrels", "100,001 to 500,000 Barrels", "500,001 to 1,000,000 Barrels", "1,000,001 to 6,000,000 Barrels", "6,000,001 Barrels and Over", "Total"))

 
```


