---
title: "Pretty_Maps"
output: html_document
---

# Making pretty maps with R and ggplot
This document uses a Tidy Tuesday dataset to set up an initial map and then steps through different packages to add features, improve layout, and customize the map.

```{r}
library(tidyverse)
library(tidytuesdayR)
# devtools::install_github("UrbanInstitute/urbnmapr")
library(urbnmapr) # County and state polygons
```

Read in the data: https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-05-11
```{r}
tuesdata <- tidytuesdayR::tt_load('2021-05-11')
broadband <- tuesdata$broadband
```
Take an initial look
```{r}
summary(broadband)
head(broadband)
```

Load urbnmapr package data, see package README for example using shape file plots (https://github.com/UrbanInstitute/urbnmapr)
```{r}
states <- urbnmapr::states
county <- left_join(countydata, counties, by = "county_fips")

# Plot states
ggplot() +
  geom_polygon(data = states, aes(x=long, y=lat, group=group), fill = "grey", color="white")

# Plot counties and states
ggplot() +
  geom_polygon(data = county, aes(x=long, y=lat, group = group), fill = "grey", color="white",size = 0.1) + # plot counties
  geom_polygon(data = states, aes(x=long, y=lat, group=group), fill = NA, color="white", size = 0.3) # Add thicker state divisions
```

Merge broadband and county data
```{r}
# Rename broadband columns so county name column named consistently in both dataframes
colnames(broadband) <- c("ST", "county_fips", "NAMELSAD", "Broadband_Avail_Per_FCC", "Broadband_Usage")

county <- county %>% mutate(county_fips = as.numeric(county_fips)) # Make FIPs ID numeric

# Merge county polygon and broadband data
broadband <- full_join(broadband, county, by = "county_fips") 
head(broadband)
```

Plot broadband availability (as of 2017) data similar to Tidy Tuesday README
```{r}
broadband <- broadband %>% mutate(PercentColor = ifelse(Broadband_Avail_Per_FCC < 0.15, "<15%", ">=15%")) 
  ggplot() +
  geom_polygon(data = broadband, aes(x=long, y=lat, group = group, fill = factor(PercentColor)), color="white",size = 0.1) + # plot counties
  geom_polygon(data = states, aes(x=long, y=lat, group=group), fill = NA, color="white", size = 0.3) + # Add thicker state divisions
    scale_color_manual(aesthetics = "fill", values = c("navy", "grey"), name = "% of people with access \n to 25Mbps/3 Mbps ", na.value = "grey36") +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    theme(panel.background = element_rect(fill="white"), axis.title.x = element_blank(), axis.title.y = element_blank(), axis.ticks.x = element_blank(), axis.ticks.y = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())
```
In most locations more than 15% of people have access to 25Mbps speed internet as of 2017.


Plot broadband use data (as of 2019) as in Tidy Tuesday README
```{r}
broadband <- broadband %>% mutate(PercentColor = ifelse(Broadband_Usage < 0.15, "<15%", ">=15%")) 
  ggplot() +
  geom_polygon(data = broadband, aes(x=long, y=lat, group = group, fill = factor(PercentColor)), color="white",size = 0.1) + # plot counties
  geom_polygon(data = states, aes(x=long, y=lat, group=group), fill = NA, color="white", size = 0.3) + # Add thicker state divisions
    scale_color_manual(aesthetics = "fill", values = c("navy", "grey"), name = "% of people using the internet \n at 25Mbps or above", na.value = "grey36") +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    theme(panel.background = element_rect(fill="white"), axis.title.x = element_blank(), axis.title.y = element_blank(), axis.ticks.x = element_blank(), axis.ticks.y = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())
```
However, it appears that as of 2019 more people with access do not use this speed of internet.

# Try changing the formatting
Here is a list of packages to consider using for inspiration: https://insileco.github.io/wiki/rgraphpkgs/#Listofpackages

Try using leafletR to create interactive map (show actual % useage when hovering over county)


Explore micromap


Explore interactive graphics packages

Explore heatmaply/plotly

To make it pretty: ggsn, ggsci, ggThemeAssist

Look at tikzDevice













