---
title: "Waffle_Tutorial"
output: html_document
---

This document walks through a simple example to make a waffle plot using starwars data from the `dplyr` package

# Setup work space
Load the `waffle` package
```{r}
# devtools::install_git("https://git.rud.is/hrbrmstr/waffle.git")
library(waffle)
```
Load other packages
```{r}
library(tidyverse)
```
Load the starwars data
```{r}
starwars <- dplyr::starwars
starwars
```

# Make an ugly waffle plot with `ggplot()`
```{r}
# The data
starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex)
# The plot
starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>%
  ggplot(aes(fill = sex, values=n)) +
    geom_waffle(nrows = 10)
```
# Improve formatting
```{r}
# The data arranged by count
starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>% arrange(., desc(n))

# The plot
starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>% arrange(., desc(n)) %>%
  ggplot(aes(fill = sex, values=n)) +
    geom_waffle(color = "white", nrows = 10) +
    theme_minimal() +
    theme_enhance_waffle()
```

Need to remove "NA"s
```{r}
newWaffle <- starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>% 
  arrange(., desc(n)) %>% 
  drop_na() %>%
  ggplot(aes(fill = sex, values=n)) +
    geom_waffle(color = "white", nrows = 10) +
    theme_minimal() +
    theme_enhance_waffle()
newWaffle
```

# Try using the `waffle()` function
```{r}
data <- starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>%
  arrange(., desc(n)) %>% 
  drop_na()

# use ggplot_build() to find the default fill colors used in newWaffle
# ggplot_build(newWaffle)$data

waffle(data, rows=10, size = 1, colors = c("#00BFC4", "#F8766D", "#C77CFF", "#7CAE00"))
```

# Turn ggplot into a pictogram
First time users may encounter the following error:

"Error: Font [Font Awesome 5 Free Solid] not found. Please install it and use extrafont to make it available to R"

To fix this error:

* Install Font Awesome 5 Free Solid, "easy" steps here: https://www.listendata.com/2019/06/create-infographics-with-r.html
* Install `emojifont` package instead (https://cran.r-project.org/web/packages/emojifont/vignettes/emojifont.html), if this doesn't work:
    + May not have correct name for emoji
    + May not work for Macs/version of R


To install Font Awesome:
```{r}
# 1
library(extrafont)

# 2 Install files

# 3 Import font from files
extrafont::font_import(path = "/Users/ahart2/Downloads", pattern = "fa-", prompt = FALSE)

# 4 load fonts
loadfonts(device="pdf")

# 5 Check
extrafont::fonttable() %>% 
  dplyr::as_tibble() %>% 
  dplyr::filter(grepl("Awesom", FamilyName)) %>% 
  select(FamilyName, FontName, fontfile)
```

Generate the pictogram using `geom_pictogram()`
```{r}
starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>% 
  arrange(., desc(n)) %>% 
  drop_na() %>%
  ggplot() +
    geom_pictogram(aes(label = sex, values=n, color = sex), n_rows = 10, make_proportional = TRUE, size=5.) +
    scale_color_manual( # Without this function call, 2 legends created
      name = NULL,
      values = c("#00BFC4", "#F8766D", "#C77CFF", "#7CAE00"),
      labels = c("male", "female", "none", "hermaphroditic")
    ) +
    scale_label_pictogram(
      name = NULL, 
      values = c("user", "user-tie", "user-ninja", "user-md"), # This is the pictogram to use
      labels = c("male", "female", "none", "hermaphroditic")
    ) +
    theme_minimal() +
    theme_enhance_waffle()
```

The colors in the legend match the data, but the colors/icons in the plot do not! Be careful with labels!
```{r}
starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>% 
  arrange(., desc(n)) %>% 
  drop_na() %>%
  ggplot() +
    geom_pictogram(aes(values = n, label = sex, colour = sex),
                   n_rows = 10, make_proportional = TRUE, size=5.) +
    scale_label_pictogram(
      name = NULL, 
      values = c("user", "user-tie", "user-ninja", "user-md"), # These are the pictograms to use
    ) +
    scale_color_manual( # Without this function call, 2 legends created
      name = NULL,
      values = c("#F8766D", "#7CAE00", "#00BFC4", "#C77CFF") # Notice I removed labels
    ) + 
    theme_minimal() +
    theme_enhance_waffle()
```

Use a uniform icon in the pictogram
```{r}
starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>% 
  arrange(., desc(n)) %>% 
  drop_na() %>%
  ggplot() +
    geom_pictogram(aes(values = n, label = sex, colour = sex),
                   n_rows = 10, make_proportional = TRUE, size=5.) +
    scale_label_pictogram(
      name = NULL, 
      values = c("male", "male", "male", "male"), # This is the pictogram to use
    ) +
    scale_color_manual( # Without this function call, 2 legends created
      name = NULL,
      values = c("#F8766D", "#7CAE00", "#00BFC4", "#C77CFF")
    ) + 
    theme_minimal() +
    theme_enhance_waffle()
```

May also use `waffle()` to create a pictogram
```{r}
data <- starwars %>% 
  filter("Attack of the Clones" %in% films) %>%
  count(sex) %>%
  arrange(., desc(n)) %>% 
  drop_na()

waffle(data, rows=10, colors = c("#00BFC4", "#F8766D", "#C77CFF", "#7CAE00"), glyph_size = 4, use_glyph = "male") # glyph_size changes the image size!
```

# Facet by birth year and show sex ratio
```{r}
starwars %>%
  mutate(year_groups = cut(birth_year, breaks = c(seq(0,900, by=50)))) %>% # Group birth years
  count(year_groups,sex) %>%
  arrange(., desc(n)) %>% 
  drop_na() %>%
  ggplot(aes(fill=sex, values=n))+
    geom_waffle(color = "white", n_rows=5, flip=TRUE) + # Flip and change # rows
    facet_wrap(~year_groups, nrow = 1, strip.position = "bottom") +
    scale_y_continuous(labels = function(x) x*5, expand=c(0,0)) + # number is same as n_rows
    scale_x_discrete() + # Don't show x-axis scale
    theme_minimal() +
    theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  labs(
    x = "Birth Year",
    y = "Count"
  )

```

Not all characters have ages so these counts will look different than previous plots.

# Lessons learned
1) Be careful with pictogram labels, often best to start with a waffle plot so you know what it should look like.

2) Waffle plots with too many variables can be hard to read
```{r}
starwars %>%
  count(eye_color) %>% # Count number in each color category
  ggplot(aes(fill = eye_color, values = n)) +
    geom_waffle()
```

3) Even if you clean up their formatting they may still be ugly
```{r}
# Use eye-colors in plot appearance
starwars %>%
  count(eye_color) %>% # Count number in each color category
  ggplot(aes(fill = eye_color, values = n)) +  # n = counts
    geom_waffle(n_rows=10, size=0.33, color="white") + 
    scale_fill_manual(values = c("black", "dodgerblue", "cadetblue3", "chocolate3", "chocolate4", "gold", "greenyellow", "brown", "darkorange", "pink", "red", "violetred", "wheat", "ghostwhite", "yellow2")) +
    theme_minimal() +
    theme_enhance_waffle()
```

4) Faceting for variables with many possible values can result in unreadable charts
```{r}
starwars %>% 
  count(homeworld, species) %>%
  ggplot(aes(fill = species, values=n)) +
  geom_waffle(color = "white", nrows = 10) +
  facet_wrap(~homeworld, nrow = 1)
```

# Helpful resources
* README file for waffle package: https://github.com/hrbrmstr/waffle
* Extra waffle and pictogram examples: https://nsaunders.wordpress.com/2017/09/08/infographic-style-charts-using-the-r-waffle-package/
* Steps to install Font Awesome for pictograms: https://www.listendata.com/2019/06/create-infographics-with-r.html
* Find icons for pictograms: https://fontawesome.com/icons?d=gallery&p=2
